\section{Introduction}\label{intro}
\subsection{Motivation}\label{intro:motivation}

With billions of IoT devices, and the prediction of billions more to be connected every year,~\cite{gartner_billions_iot, iot_forecast_update} practically every problem in IoT is realised at large scale. The connection of so many devices typically requires, or at least results in, manufacturers' rapid iteration of software and continual deployment of updates - in order to remain competitive, and to respond to the large number of bugs uncovered by such a user base.~\cite{cd_for_iot,iot_software_release} This iterative development and frequent deployment may lead easily to unintended incompatibilities between server APIs,~\cite{client_server_coupling} and those outdated or mis-configured APIs that client devices attempt to use. Additionally, many of the devices expected to be deployed will be deeply embedded in their operating context: light switches, concrete-borne traffic or structural sensors in roads or buildings,~\cite{construction_and_iot, smart_concrete} et cetera. Such devices have relatively long lifetimes and may be offline for long periods or more easily forgotten about entirely compared to, say, an IoT teapot in a consumer's kitchen.

Client-originating communication with a server that results in an error may be of great concern to the end-user: important structural integrity readings not being delivered, or a teabag supplier not notified when the teapot's brewed through a batch - but equally might present only to the manufacturer, or an owner separate to the concerned `user' in cases such as traffic and structural sensors.

With the fairly common separation~\cite{backend_for_frontend, front_backend_keep_separate, front_backend_separation_of_concerns} of software engineers developing back-end servers and those working on the clients or IoT devices with which they will communicate,~\cite{fielding_rest} it is easy to imagine a scenario in which a small but important detail is miscommunicated: is the URL path `\code{/foobar}' or `\code{/foobars}'; is the data returned in a flat JSON object, or is it nested under the key `\code{data}', perhaps with pagination? These are questions that the third-party or hobbyist developer, typically working against an existent system, is able to answer by manual inspection. The team working in tandem with the creation of the API, however, is reliant on communication of these details;~\cite{team_comms_silo} both at first instance, and if there should be any later change. Even in the presence of good testing,~\cite{front_backend_testing} such bugs can slip through.

Moreover, if requirements do change and the API is affected, then even if it is communicated to the team developing the client device it may be a significant undertaking to transform the data sent or modify the URLs in many places. While the bulk may be a straightforward pattern replacement, the potential for a vital few~\cite{pareto_software_dev} cases not matching the pattern necessitates a considered study, or extremely good test coverage.

\input{introduction/requirements}