\subsection{Requirements} \label{intro:req}

To obviate uncertainty surrounding the correctness of client devices' consumption of an API, we would like a system that leverages the type system and compiled nature of a language used for embedded systems, in order to provide certain validations at compile-time.

There are many ways in which interaction between a client and server can fail; indeed at all layers in the TCP/IP stack.~\cite{networks_distributed_systems} It is the application layer that is of concern here, wherein the failure modes of HTTP (or CoAP\footnote{CoAP is a stripped-down sibling of HTTP's, designed specifically for constrained IoT devices and networks.~\cite{rfc7252}}) are described by the \code{4xx} and \code{5xx} status code ranges,~\cite{rfc7231,rfc7252} representing client and server errors respectively.

\subsubsection{Preventable client errors} \label{intro:req:preventable}
HTTP's client error status codes include `\code{405 Method Not Allowed}', indicating that the method used (\code{GET}, \code{POST}, et al.) is not an action available for the resource on which it was attempted, and `\code{400 Bad Request}', which is typically indicative of a malformed request body - such as \mintinline{json}`{"id": 42}` when the field \mintinline{json}{"id"} is expected to contain a string. Server error codes mostly pertain to ephemeral issues - unavailability due to the machine rebooting, or the web server having collapsed under sudden demand, for example. It is the possibility of client errors, the \code{4xx} range, that we can aim to avoid. \cite{rfc7231}

Thus, the system should aim to prevent the possibility, so far as possible, of errors occurring due to bugs in the client device code. \Cref{tbl:req:preventable-errors} details the codes it is expected to be possible to prevent.

\begin{spacing}{1.1}
\begin{table}[!ht]
\caption{Client errors anticipated to be compile-time avoidable}
\label{tbl:req:preventable-errors}
\centering
\begin{adjustwidth}{-1in}{-1in}
\begin{tabular}{lll}
\toprule
\multicolumn{1}{c}{\multirow{2}{*}{\textbf{Status Code}}} & \multicolumn{2}{c}{\textbf{Example}}
\\\cmidrule(lr){2-3}
& \multicolumn{1}{c}{\textbf{Use}} & \multicolumn{1}{c}{\textbf{Problem}}
\\\midrule
400 Bad Request & Request body has structure $A$ & Expected structure $B \neq A$
\\\cmidrule(lr){1-3}
401 Unauthorized & Unauthenticated request & Endpoint requires authentication
\\\cmidrule(lr){1-3}
404 Not Found & \code{GET /foo} & \code{/foo} does not exist
\\\cmidrule(lr){1-3}
405 Method Not Allowed & \code{PATCH /foo} & \code{/foo} does not support \code{PATCH}
\\\cmidrule(lr){1-3}
406 Not Acceptable & \code{Accept: application/json} & Endpoint does not provide JSON
\\\cmidrule(lr){1-3}
407 Proxy Auth. Required & Unauthenticated proxied request & Endpoint requires authentication
\\\cmidrule(lr){1-3}
411 Length Required & No \code{Content-Length} header & Endpoint requires the header
\\\cmidrule(lr){1-3}
412 Precondition Failed & \code{PATCH /foo \{"state": 3\}} & Endpoint cannot transition  $1 \to 3$
\\\cmidrule(lr){1-3}
415 Unsupported Media Type & \code{POST /foo \{\}} & Endpoint does not accept JSON
\\\cmidrule(lr){1-3}
417 Expectation Failed & \code{Expect: 100-continue} & Endpoint cannot respond \code{100}
\\\cmidrule(lr){1-3}
421 Misdirected Request & Connection to $A$ reused against $B$ & $B$ has greater security requirements
\\\cmidrule(lr){1-3}
424 Failed Dependency & \code{POST /dependant \{"ref": "foo"\}} & Dependency must be created first
\\\cmidrule(lr){1-3}
426 Upgrade Required & \code{HTTP/1.0 ...} & Endpoint requires \code{HTTP/1.1} 
\\\cmidrule(lr){1-3}
428 Precondition Required & \code{PUT ...} & Must \code{If-Match} on a priori state
\\\bottomrule
\end{tabular}
\end{adjustwidth}
\end{table}
\end{spacing}

\FloatBarrier
\Cref{tbl:req:preventable-errors} forms a proper subset of possible client errors, since some such as \code{410 Gone} (which indicates a resource did exist, but no longer does), \code{403 Forbidden} (user has correctly authenticated, but is not authorised for this action), and \code{409 Conflict} (modification to the resource is in conflict with its a priori state) are caused by some aspect variable in the lifetime of the application.

Further, codes such as \code{409} and \code{410} could be due to the API interaction of other actors, if multiple devices have access to the same resource, in which case there is no hope to prevent the error entirely.

Others may be possible to prevent partially - such as the \code{422 Un\-proc\-ess\-able Entity} which typically signals a validation error within a (structurally sound, so \code{400} is inappropriate) request body, and those pertaining to data in the request being too long. \cite{rfc7252}

\subsubsection{Sessions} \label{intro:req:sessions}
Preventing some of the HTTP status codes listed in \cref{tbl:req:preventable-errors} will require knowledge of a `session' of multiple requests: validating that an authentication request has occurred prior to one that requires it, for example.

More generally, those in~\cref{tbl:req:preventable-errors} are the errors that would not be so under different circumstances: the request is syntactically correct, but its semantics are not achievable (or not allowable) on the particular occasion.

Thus, for maximum reduction in possibility of run-time HTTP errors, a solution will be required to analyse the state over an interaction of a series of requests, in addition to its analysis of each request in itself.

\subsubsection{Unpreventable client \& server errors} \label{intro:req:unpreventable}
Although we cannot hope through any amount of analysis of client software to stop server errors from occurring, it would be desirable to make the programmer aware of the possibility; to force him, via the compiler, to handle the error cases in a considered manner.

\subsubsection{Productivity} \label{intro:req:productivity}
A `softer', but no less important requirement is that the implementation allow the user (viz. the programmer of the client device software) to be productive. An implementation that burdens the user with additional work and without perceivable gain simply would not be used; even if another party, perhaps the end-user of the device, would benefit.

Throughout the industry, burdensome technical advantage is traded off against ease of development; the latter almost always wins. The result is that formally verified programs are mostly confined to research, and the most safety-critical corners of the defence and aviation industries.~\cite{formal_verification_vs_testing} Even compiled languages seem increasingly to be used only where necessary, with efforts to bring higher-level languages such as JavaScript to desktop platforms,~\cite{electron} and even to embedded~\cite{jerryscript} devices. Statically typed languages form a cross-section that seems somewhat resistant to this trend,~\cite{static_vs_dynamic_productivity} with projects such as the popular TypeScript~\cite{typescript} that transpiles to JS offering the boost to productivity that comes with the earlier warning of a large category of errors.

Thus, the implementation presented here should be an aid, not a hindrance, to the user. It should make handling the failure modes discussed in \cref{intro:req:unpreventable} straightforward, be transparent about why certain options are unavailable, and, in accordance with \cref{intro:motivation}, be of benefit to the development of the client device without impacting or requiring changes on any server with which it communicates.